Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-19 main.tex
Added citations to lme4 and R.

Before fitting the model, we scale each independent variable \(x_{ij}^k\) such
that they have a mean of zero and a standard deviation of one by cluster-mean
centering, as recommended by \cite{Enders2007}, with clusters being an
individual subject. The clusters are chosen as all data from an individual
subject because we are only interested in the association between the state of
the balance-assist system and the outcome of the perturbation. We expected
there to be a variation between participants in how well they are able to
resist the perturbation. However, this was not true. Cluster-mean centering
shows there to be no variation between participants. This fact allows us to
utilize a simple single-level logistic regression model, instead of a
multilevel model. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(g=10\) trials and one
for the 10~\si{\kph}, \(g=8\) trials and we evaluate our hypothesis for each
set of data.

@@

Turning the balance-assist system on significantly \((p<0.05)\) reduces the
odds that a perturbation results in a fall while cycling at a speed of
6~\si{\kph}. Figure~\ref{fig:probability-6kph} visualizes the probability of
falling as a function of the mean and centered angular impulse per participant
for the balance assist state on and off while keeping all other explanatory
factors at their centered mean values. This figure is created by setting all
explanatory variables to there mean and calculating the probability from
Eq~\ref{eq:log-regress} for only change in angular impulse given the estimates
in Table~\ref{tab:freq-coefs-6}. The table indicates that the balance-assist
system halves (0.53) the odds that a perturbation results in a fall. This
figure shows that for relatively large impulses the probability to fall is
unity for both states of balance assist on and off. And for relatively small
impulses the probability to fall is null for both states. But for impulses in
the magnitude region (-1 to 0.5 STD), i.e. around the mean-centered fall
threshold, the probability of falling is significantly lowered with the balance
assist system on. The skewness of the probability curves arrives from the
interaction effects.\todo{check this statement about skewedness this and think
about it} Figure~\ref{fig:probability-10kph} shows the same result for the
10~\si{\kph} trials which has a similar trend of reducing the probability to
fall with the balance assist system turned on, but the effect is not
significant.
%
\begin{figure}
  \centering
  % NOTE : We could uncomment the longer subcaption if titles removed from
  % figures. It is currently redundant.
  \subcaptionbox{%Fall probability for the 6~\si{\kph} trials.
    \label{fig:probability-6kph}}{
    \includegraphics[width=80mm]{figures/predicted_fall_probability_6kmh.png}
  }
  \subcaptionbox{%Fall probability for the 10~\si{\kph} trials.
    \label{fig:probability-10kph}}{
    \includegraphics[width=80mm]{figures/predicted_fall_probability_10kmh.png}
  }
  \caption{Comparison of predicted fall probability for balance-assist on
    (blue) or off (red) when all other predictor variables are fixed at zero,
    except for the interaction effect between the balance-assist and the
    angular impulse. The ordinate axis is standard deviation around the mean
    (all perturbations), centered per participant.}
  \label{fig:probability}
\end{figure}

>>>>

Before fitting the model, we scale each independent variable \(x_{ij}^k\) such
that they have a mean of zero and a standard deviation of one by cluster-mean
centering, as recommended by \cite{Enders2007}, with clusters being an
individual subject. The clusters are chosen as all data from an individual
subject because we are only interested in the association between the state of
the balance-assist system and the outcome of the perturbation. We expected
there to be a variation between participants in how well they are able to
resist the perturbation. However, this was not true. Cluster-mean centering
shows there to be no variation between participants. This fact allows us to
utilize a simple single-level logistic regression model, instead of a
multilevel model. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(g=10\) trials and one
for the 10~\si{\kph}, \(g=8\) trials and we evaluate our hypothesis for each
set of data. Statistics were computed with the lme4 version
1.1~\citep{Bates2015} package in R version 4.1.3~\citep{RCoreTeam2021}.

@@

Turning the balance-assist system on significantly \((p<0.05)\) reduces the
odds that a perturbation results in a fall while cycling at a speed of
6~\si{\kph}. Figure~\ref{fig:probability-6kph} visualizes the probability of
falling as a function of the mean and centered angular impulse per participant
for the balance assist state on and off while keeping all other explanatory
factors at their centered mean values. This figure is created by setting all
explanatory variables to there mean and calculating the probability from
Eq~\ref{eq:log-regress} for only change in angular impulse given the estimates
in Table~\ref{tab:freq-coefs-6}. The table indicates that the balance-assist
system halves (0.53) the odds that a perturbation results in a fall. This
figure shows that for relatively large impulses the probability to fall is
unity for both states of balance assist on and off. And for relatively small
impulses the probability to fall is null for both states. But for impulses in
the magnitude region (-1 to 0.5 STD), i.e. around the mean-centered fall
threshold, the probability of falling is significantly lowered with the balance
assist system on. The skewness of the probability curves arrives from the
interaction effects.\todo{check this statement about skewedness this and think
about it} Figure~\ref{fig:probability-10kph} shows the same result for the
10~\si{\kph} trials which has a similar trend of reducing the probability to
fall with the balance assist system turned on, but the effect is not
significant.
%
\begin{figure}
  \centering
  % NOTE : We could uncomment the longer subcaption if titles removed from
  % figures. It is currently redundant.
  \subcaptionbox{Fall probability for the 6~\si{\kph} trials.
    \label{fig:probability-6kph}}{
    \includegraphics[width=80mm]{figures/predicted_fall_probability_6kmh.png}
  }
  \subcaptionbox{Fall probability for the 10~\si{\kph} trials.
    \label{fig:probability-10kph}}{
    \includegraphics[width=80mm]{figures/predicted_fall_probability_10kmh.png}
  }
  \caption{Comparison of predicted fall probability for balance-assist on
    (solid) or off (dashed) when all other predictor variables are fixed at zero,
    except for the interaction effect between the balance-assist and the
    angular impulse. The abscissa is standard deviation around the mean (all
    perturbations), centered per participant.}
  \label{fig:probability}
\end{figure}
